Data Specialization for a Staged Lambda Calculus

### 2014-11-07 paper/splitting.tex
Rough editing pass through the stage 1 splitting description.

The splitting algorithm for stage~\bbone\ terms, as specified by the
judgment $e \splitonesym [c,l.r]$ in \cref{fig:termSplitOne}, proceeds
recursively on the structure of~$e$.
When $e$ is a terminal (a variables or base constant)
splitting yields a combined result formed by tupling $e$ with a \texttt{()} precomputation, and the trivial resumer \texttt{()}. (Stage~\bbone\ terminals, by definition, contain no stage~\bbtwo\ content.)  For example, \texttt{3} splits into combined result \texttt{(3,())} and resumer \texttt{\_=>()}.

@@

When $e$ is a non-terminal (except $\next$),
splitting descends into $e$, recursively splitting subterms
and generating combined terms $c_1,\ldots,c_n$.
The new combined term of $e$ then works by binding $c_1,\ldots,c_n$
to the patterns $(y_1,z_1),\ldots,(y_n,z_n)$
to isolate their immediate results and precomputations.
The immediate result of $e$ is then formed by replacing its subterms with $y_1,\ldots,y_n$, 
and the precomputation is formed from the tuple of $z_1,\ldots,z_n$.

At {\tt case}s, only one of the branches' combined results is executed, and its precomputation injected and bundled with the precomputation of the predicate. 
This precomputation is then cased in the resumer to resume the stage \bbtwo\ portion of the correct branch.
This rule and $\pause$ are the only two places where nontrivial information is added to the precomputation.

[FUNCTIONS]

>>>>

The splitting algorithm for stage~\bbone\ terms, as specified by the
judgment $e \splitonesym [c,l.r]$ in \cref{fig:termSplitOne}, proceeds
recursively on the structure of~$e$.
When $e$ is a terminal (a variable or base constant)
splitting yields a combined result formed by tupling $e$ with a \texttt{()} precomputation, and the trivial resumer \texttt{()}. (Stage~\bbone\ terminals, by definition, contain no stage~\bbtwo\ content.)  For example, \texttt{3} splits into the combined result \texttt{(3,())} and resumer \texttt{\_=>()}.

@@

For all non-terminals (except $\next$),
splitting descends into $e$, recursively splitting its $n$ subterms
to produce their respective combined terms $c_1,\ldots,c_n$ and resumers $r_1, \ldots, r_n$.
The combined term of $e$ is formed by binding $c_1,\ldots,c_n$
to the patterns $(y_1,z_1),\ldots,(y_n,z_n)$
to isolate immediate results from the results of precomputations. Then,
the immediate result of $e$ is formed by replacing $e$'s subterms with $y_1,\ldots,y_n$.
The resumer binds the precomputation results $z_1,\ldots,z_n$ to an
argument $(l_1,\ldots,l_n)$ in a term that has the same structure
of~$e$ but where each subterm is replaced by its resumer ($r_i$'s).

The combined term of a {\tt case} executes one of the branches' combined terms based on the immediate result $y_1$ of the predicate.
The precomputation result $z_i$ for this branch is injected and bundled with that of the predicate ($z_1$).   
$z_i$ is cased in the resumer to determine which branch's resumer should be executed.
{\tt case} and $\pause$ are the only two rules where splitting adds non-trivial logic is added to the precomputation.

% We call attention to the rule for {\tt case}s since it, and the rule for $\pause$, where splitting adds non-trivial logic to the precomputation.

Function introduction has a \texttt{()} precomputation result,
since functions are already fully reduced in our semantics.
However, since the body of a function may itself be multi-stage, splitting must continue into it.
The immediate result is a new function formed from the stage~\bbone\ part of the original body.
The resumer is a new function formed out of the stage \bbtwo\ part of the original body.
It is the responsibility of the application site to save the precomputation of the function body
and pass it to the resumer version of the function.
